---
title: "Comparison of Position Weight Matricies using Euclidean Distance"
output:
  word_document: default
  html_notebook: default
  html_document: default
---

```{r}
library(ggplot2)
library(readr)
library(dplyr)
library(stringr)
library(cowplot) #plot_grid
library(readxl) #load Jadwin data excel_sheets, read_excel

source("./functions.R")
```

Set before and after parameters for # of aa
```{r}
#Number of Amino Acids to include Before and After Tyrosine
before<-1
after<-4

#Remove central tyrosine to avoid biasing distance values
remove_Y<-TRUE
```

Load Jadwin data
```{r}
JadwinFile<-"./LOGO data.xlsx"
sheetNames<-excel_sheets(JadwinFile)
for (x in sheetNames){
  tempData<-read_excel(JadwinFile, sheet = x)
  print(x)
  if ("Weighted LIst" %in% colnames(tempData)){
    tempData$`Weighted List`<-tempData$`Weighted LIst`
  }
  assign(x, tempData$`Weighted List`)
}
```

Shorten Jadwin data to parameters listed above, filter out sequences with too few aa & remove Y from sequences
```{r}
`CRK Jadwin et al`<-substr(`CRK Jadwin et al`,(7-before),(7+after))
`GRB2 Jadwin et al`<-substr(`GRB2 Jadwin et al`,(7-before),(7+after))
#filter
`CRK Jadwin et al`<-`CRK Jadwin et al`[(nchar(`CRK Jadwin et al`)>(before+after))]
`GRB2 Jadwin et al`<-`GRB2 Jadwin et al`[(nchar(`GRB2 Jadwin et al`)>(before+after))]

if (remove_Y){
  `CRK Jadwin et al`<-paste(substr(`CRK Jadwin et al`,1,before),substr(`CRK Jadwin et al`,before + 2,before+after+1),sep = "")
  `GRB2 Jadwin et al`<-paste(substr(`GRB2 Jadwin et al`,1,before),substr(`GRB2 Jadwin et al`,before + 2,before+after+1),sep = "")
}
```

Create PWM lists for Jadwin data
```{r}
  jadwin_list<-c("CRK Jadwin et al","GRB2 Jadwin et al")
  jadwin_pwm_list<-lapply(jadwin_list, function(x){
  PWM(get(x),do.pseudocounts = T)
})
```

Functions - load and modify Tinti files
Filters for aa sequences of correct length
Removes Y from sequences
```{r, message=FALSE, warning=FALSE}
#read tinti data
filenames<-list.files("./normalization/SH2_all_noreplicates_02-05-2007/",full.names = TRUE)
tinti_data<-lapply(filenames,read_csv)
tinti_data<-do.call(rbind,tinti_data)
#clean data
tinti_data<-
  tinti_data %>%
  filter(spot_flag != "BAD") %>%
  filter(peptide_sequence != "none") #%>%
#trim
tinti_data$peptide_sequence<-substr(tinti_data$peptide_sequence,7-before,after+7)
#filter
tinti_data<-tinti_data%>%filter(nchar(peptide_sequence)>(before+after))
if (remove_Y){
  tinti_data$peptide_sequence<-paste(substr(tinti_data$peptide_sequence,1,before),substr(tinti_data$peptide_sequence,before + 2,before+after+1),sep = "")
}
```


Calculate standard deviation, median and threshold samples
Applies thresholding used in Tinti paper
```{r}
statistics<-tinti_data%>%group_by(domain_shortlabel)%>%summarise(stddev = sd(log_ratio),med = median(log_ratio))
tinti_data<-left_join(tinti_data,statistics,by="domain_shortlabel")
tinti_data<-tinti_data%>%filter(log_ratio>(2*stddev+med))
```

Creating PWMs for tinti data
```{r}
sh2_list<-levels(as.factor(tinti_data$domain_shortlabel))
PWM_list<-lapply(sh2_list, function(x) {
  sequences<-tinti_data%>%filter(domain_shortlabel ==x)
  PWM(sequences$peptide_sequence,do.pseudocounts = T)
})
```


Calculate distances for all Jadwin-Tinti SH2 combinations and arrange into dataframe
```{r}
#creates empty data frame for distances
enhanced_df<-data.frame(pair = c(), euclidean = c())

#iterates through all combinations of jadwin and tinti pwms
for (i in 1:(length(sh2_list))){
  for (j in 1:length(jadwin_list)){
    #creates temp dataframe with row containing name of tinti and jadwin samples and caculated distance for that pair
    temp<-data.frame(pair1=sh2_list[i],pair2 = jadwin_list[j],euclidean=get_distance(PWM_list[[i]],jadwin_pwm_list[[j]]),denom =(dim(PWM_list[[i]])[2]+dim(jadwin_pwm_list[[j]])[2]))
    #stores temporary row in enhanced_df dataframe
    enhanced_df<-rbind(enhanced_df,temp)
    #saves copy of CRK(GCG) tinti-jadwin distance
    if ((sh2_list[i] == "CRK") & (jadwin_list[j] == "CRK Jadwin et al")){
      CRK_euclidean<-temp[3]$euclidean
    }
    #saves copy of GRB2 tinti-jadwin distance
    if ((sh2_list[i] == "GRB2 ASH") & (jadwin_list[j] == "GRB2 Jadwin et al")){
      GRB2_euclidean<-temp[3]$euclidean
    }
  } 
}
```

Set bin size for final plots
```{r}
binsize=0.1
```

Plot GRB2-TintiGRB2 distance vs all GRB2-Tinti distances
```{r}
plotData<-enhanced_df%>%filter(pair2=="GRB2 Jadwin et al")
#creates dataframe with names and distances for the two pairs which correspond to the jadwin data(ie. Grb2-Grb2 and GCG-Crk)
lines_df<-data.frame(name=c("GRB2_euclidean","CRK_euclidean"),euclidean=c(GRB2_euclidean,CRK_euclidean),grp=letters[3:4])
lines_df<-lines_df%>%filter(name == "GRB2_euclidean")
plot1<-create_histogram(plotData, binsize, lines_df)
```

Plot GCG-TintiCrk distance vs all GCG-Tinti distances
```{r}
plotData<-enhanced_df%>%filter(pair2=="CRK Jadwin et al")
lines_df<-data.frame(name=c("GRB2_euclidean","CRK_euclidean"),euclidean=c(GRB2_euclidean,CRK_euclidean),grp=letters[3:4])
lines_df<-lines_df%>%filter(name == "CRK_euclidean")
plot2<-create_histogram(plotData, binsize, lines_df)
```

Combine Plots into single figure
```{r}
title <- ggdraw() + draw_label("Distribution of Euclidean Distances", fontface='bold')
p<-plot_grid(plot1, plot2, align='h', labels=c('A', 'B'))
pg<-plot_grid(title, p, ncol=1, rel_heights=c(0.1, 1)) # rel_heights values control title margins
save_plot("pg.pdf", pg, base_aspect_ratio = 1.5)
pg
```



Find sh2 domains which are closer to Jadwin GCG than tinti GCG
```{r}
GCGBar<-enhanced_df%>%filter(pair2=="CRK Jadwin et al")%>%arrange(euclidean)
create_bar_graph(GCGBar,"CRK")
```
Calculate number for figure
```{r}
lines_df<-data.frame(name=c("GRB2_euclidean","CRK_euclidean"),euclidean=c(GRB2_euclidean,CRK_euclidean),grp=letters[3:4])
calculated_percentage_lower(enhanced_df,"CRK Jadwin et al", "CRK_euclidean",lines_df)
```

Find sh2 domains which are closer to Jadwin GRB2 than tinti GRB2
```{r}
GRB2Bar<-enhanced_df%>%filter(pair2=="GRB2 Jadwin et al")%>%arrange(euclidean)
create_bar_graph(GRB2Bar,"GRB2 ASH")
```
Calculate number for figure
```{r}
calculated_percentage_lower(enhanced_df,"GRB2 Jadwin et al", "GRB2_euclidean",lines_df)
```

Data used for CRK/GCG figures
```{r}
enhanced_df%>%filter(pair2=="CRK Jadwin et al")
```

Data used for GRB2 figures
```{r}
enhanced_df%>%filter(pair2 == "GRB2 Jadwin et al")
```

Write session information to document packages
```{r}
writeLines(capture.output(sessionInfo()), "session_info.txt")
```
